---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Avery Driscoll"
date: "2-7-2022"
output: html_document
---

```{r setup, include = FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```

# Assignment

## Question 1

Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r, warning = F, message=F}
#set and read the site URL
site_url <- 'https://snowstudies.org/archived-data/'
webpage <- read_html(site_url)

#Extract the links for the SASP forcing and SBSP forcing dataset
links <- webpage %>% #go to the webpage
  html_nodes('a') %>% #get all the links on the webpage 
  .[grepl('forcing',.)] %>% #only take links that include the string "forcing" 
  html_attr('href') #clean out html text to just keep the links

```

## Question 2 
Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r, warning = F, message=F}
#This breaks the link strings in 8 separate strings, each of which contains the text between the forward slashes, and stores them in a character matrix. The 8th column of the matrix is selected, which contains just the filenames from the urls. The text "data/" is added to the start of the strings so that the downloaded data will be automatically stored in the data folder.
file_names <- paste0('data/', str_split_fixed(links,'/',8)[,8])

#This applies the download.file function to the links and file_names vector, which downloads the datasets from the links vector and names them according to the file_names vector.
map2(links[1:2],file_names[1:2],download.file)

```


## Question 3 

Write a custom function to read in the data and append a site column to the data. 

```{r, warning = FALSE, message=F}

#The headers are not included in the .txt files. This code grabs the variable names from the metadata pdf file. Not all of these are column names are correct, but columns 1 (year), 2 (month), 3 (day), 4 (hour), 7 (precipitation), and 10 (air tempearature) are correct and contain all of the data we need.
library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")

#This function pulls the site name out of the file name and reads in the .txt files
read_met_data <- function(file){
  name = str_split_fixed(file,'_',2)[,2] %>% #remove text up through the first _ ("data/SBB_")
    gsub('_Forcing_Data.txt','',.) #remove the text "_Forcing_Data.txt"
  df <- read_fwf(file) %>% #This read_fwf function is able to handle fixed width text datasets
    select(year = X1, month = X2, day = X3, hour = X4, precip = X7, air_temp = X10) %>% #choose and name vars of interest based on column numbers from headers output
    mutate(site = name)#Add a "site" column containing the  name 
}

```

## Question 4

Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r, warning = FALSE, message=F}

met_data <- map_dfr(file_names, read_met_data)
summary(met_data)

```

## Question 5

Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

This code calculates annual average air temperatures for each site and plots them over time. The 2003 temperature is very low because the dataset only contains November and December measurements, so 2003 data should be excluded. A quick check to ensure that all of the months are included in the other years indicates that 2011 data is also incomplete; no months after Oct. are included. 

```{r, warning = FALSE, message=F}

#Calculated mean temp by site and year
ann_temp_data <- met_data %>%
  group_by(year,site) %>%
  summarize(mean_temp = mean(air_temp, na.rm = T))

#Plot mean temp by site and year including all years
ggplot(ann_temp_data, aes(x = as.factor(year), y = mean_temp, group = site)) + 
  geom_line(aes(color = site)) + 
  geom_point(aes(color = site)) + 
  theme_classic() + 
  xlab("Year") + ylab("Mean air temperature (K)") + 
  scale_color_manual(values = c("grey30", "dodgerblue3")) + 
  theme(legend.position = c(0.9, 0.2))

#Count the months included for each site and year to make sure other years are complete; 2011 is not
check <- met_data %>%
  group_by(year, site) %>%
  summarize(n_months = length(unique(as.character(month))))

```

This code reproduces the figure above but excludes 2003 and 2011 because of the incomplete data. In general, the mean annual air temperature is lower at the SBSP site by about 1.5 degrees. Although the average 2004 temperature for SBSP is much lower than average, the data in 2004 appear to be complete for both sites and a boxplot of the hourly values at SPSB in 2004 by month does not reveal any obvious outliers.

```{r, warning = FALSE, message=F}
#Filtering out 2003 data and replotting
ann_temp_data <- met_data %>%
  filter(!(year %in% c(2003, 2011))) %>%
  group_by(year,site) %>%
  summarize(mean_temp = mean(air_temp, na.rm = T))

#Plot annual average air temperature over time with only complete years
ggplot(ann_temp_data, aes(x = as.factor(year), y = mean_temp, group = site)) + 
  geom_line(aes(color = site)) + 
  geom_point(aes(color = site)) + 
  theme_classic() + 
  xlab("Year") + ylab("Mean air temperature (K)") + 
  scale_color_manual(values = c("grey30", "dodgerblue3")) + 
  theme(legend.position = c(0.9, 0.2))

#Boxplot of hourly 2004 values at SBSP by month just to check for obvious outliers; no concerning values obvious.
sbsp_2004 <- met_data %>% 
  filter(year == 2011, site == "SBSP")

ggplot(sbsp_2004, aes(x = as.factor(month), y = air_temp)) + 
  geom_boxplot() + xlab("Month") + ylab("Air temperature (K)") + 
  theme_classic()

```

## Question 6

Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?

There are no months in the years from 2005 to 2010 during which average air temperatures at SBSP are warmer than at SASP.

```{r, warning = FALSE, message=F}

#Create the function to calculate average temperature by site and month, then create and output a line graph
temp_func <- function(yr, df){
  #calculate the mean temperature by site and month
  plot_df <- df %>%
    filter(year == yr) %>%
    group_by(site, month) %>%
    summarize(mean_temp = mean(air_temp, na.rm = T))
  
  #plot the mean temperature by month and site
  a <- ggplot(plot_df, aes(x = as.factor(month), y = mean_temp, group = site)) + 
    geom_line(aes(color = site)) + 
    geom_point(aes(color = site)) + 
    xlab("Month") + ylab("Mean air temperature (K)") + 
    theme_classic() + labs(title = as.character(yr)) + #label the plot with the selected year
    scale_color_manual(values = c("grey30", "dodgerblue3")) + 
    theme(legend.position = c(0.9,0.9))
  
  print(a)#this is needed to display the plot
}

#Loop through 2005-2010 to produce figures
for(i in c(2005:2010)){
  temp_func(i, met_data)
}

```


## Bonus

1. Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site.

```{r, warning = F, message=F}

#Create a DOY column from the year, month, day columns
met_data$date <-  as.POSIXct(paste(met_data$year, met_data$month, met_data$day, sep = "/"), format = "%Y/%m/%d")
met_data$doy <- as.integer(strftime(met_data$date, format = "%j"))
  
#Average precip by DOY and site, and multiply by 24 to get from hourly to daily values. 
daily_avg_precip <- met_data %>%
  group_by(doy, site) %>%
  summarize(mean_pre = mean(precip, na.rm = T)) %>%
  mutate(daily_pre = mean_pre*24)

#Plot precipiation by doy
ggplot(daily_avg_precip, aes(x = doy, y = daily_pre)) + 
    geom_line(aes(color = site)) + 
    geom_point(aes(color = site)) + 
    xlab("DOY") + ylab("Mean daily precipitation (kg m-2 s-1)") + 
    theme_classic() +
    scale_color_manual(values = c("grey30", "dodgerblue3")) + 
    theme(legend.position = c(0.1,0.9))

```

2. Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 

```{r, warning = F, message=F}

#Create function to calculate precipitation by DOY for each year and make a plot
precip_func <- function(yr, df){
  plot_df <- df %>%
  filter(year == yr) %>% #subset the selected year
  group_by(doy, site) %>% 
  summarize(mean_pre = mean(precip, na.rm = T)) %>% #calculate average hourly precipitation by doy and site
  mutate(daily_pre = mean_pre*24) #multiply average hourly values by 24 to get average daily values

  #plot the mean precip by doy and site
 a <- ggplot(plot_df, aes(x = doy, y = daily_pre)) + 
    geom_line(aes(color = site)) + 
    geom_point(aes(color = site)) + 
    xlab("DOY") + ylab("Mean daily precipitation (kg m-2 s-1)") + 
    theme_classic() + labs(title = as.character(yr)) +
    scale_color_manual(values = c("grey30", "dodgerblue3"))

  print(a)#this displays the plot
}

#Loop through 2004-2010 to produce figures
for(i in c(2004:2010)){
  precip_func(i, met_data)
}

```

